---
title: "Working with  JavaScript in R"
author: "Jeroen Ooms"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Working with JavaScript in R}
  \usepackage[utf8]{inputenc} 
output:
  knitr:::html_vignette:
    toc: yes
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = "")
options(width=100L)
library(js)
```

## What is JS?

The js package implements bindings to several popular JavaScript libraries for validating, reformating, optimizing and analyzing JavaScript code. It builds on the [V8 package](http://cran.r-project.org/web/packages/V8/vignettes/v8_intro.html) (a fully standalone JavaScript engine in R) to call out to these libraries. 

## Syntax Validation

Several R packages allow the user to supply JavaScript code to be used as callback function or configuration object within a visualization or web application. By validating in R that the JavaScript code is syntactically correct and of the right type before actually inserting it in the HTML, we can avoid many annoying bugs. 

The `js_typeof` function simply calls the `typeof` operator on the given code. If the code is syntactically invalid, a SyntaxError will be raised.

```{r}
callback <- 'function(x, y){ 
  var z = x*y ;
  return z;
}'
js_typeof(callback)
```

Same for objects:

```{r}
conf <- '{
  foo : function (){},
  bar : 123
}'
js_typeof(conf)
```

Catch JavaScript typos:

```{r error=TRUE, purl = FALSE}
js_typeof('function(x,y){return x + y}}')
```

## Script Validation

A JavaScript program typically consists of script with a collection of JavaScript statements. The `js_validate_script` function can be used to validate an entire script. 

```{r}
jscode <- readLines(system.file("js/uglify.min.js", package="js"), warn = FALSE)
js_validate_script(jscode)
```

Note that JavaScript does not allow for defining anonymous functions in the global scope:

```{r}
js_validate_script('function(x, y){return x + y}', error = FALSE)
```

To validate individual functions or objects, use the `js_typeof` function. 

## Uglify: reformatting and optimization

One of the most popular and powerful libraries for working with JavaScript code is [uglify-js](https://www.npmjs.com/package/uglify-js). This package provides an extensive toolkit for manipulating the syntax tree of a piece of JavaScript code.

The `uglify_reformat` function parses a JavaScript code and then feeds it to the [uglify code generator](http://lisperator.net/uglifyjs/codegen) which converts it back to a JavaScript text, with custom formatting options. This is nice for fixing whitespace, semicolons, etc.

```{r}
code <- "function test(x, y){ x = x || 1; y = y || 1; return x*y;}"
cat(uglify_reformat(code, beautify = TRUE, indent_level = 2))
```

However the more impressive part of uglify-js is the [compressor](http://lisperator.net/uglifyjs/compress) which refactors the entire syntax tree, effectively rewriting your code into a more compact but equivalent program. The `uglify_optimize` function in R is a simple wrapper which parses JavaScript and then feeds it to the compressor.

```{r}
cat(code)
cat(uglify_optimize(code))
```

You can pass [compressor options](http://lisperator.net/uglifyjs/compress) to `uglify_optimize` to control the various uglify optimization techniques.

## JSHint: code analysis

JSHint will automatically detect errors and potential problems in JavaScript code. The `jshint` function is R will return a data frame where each row is a problem detected by the library (type, line and reason of error): 

```{r}
code <- "var foo = 123"
jshint(code)
```

JSHint has many [configuration options](http://jshint.com/docs/options/) to control which types of code propblems it will report on. 
